Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar

### 2017-05-13 deforestmap.Rmd
New bib file and intro

---
title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

@@

Separated from the African continent and the Indian plate about 165 and 85 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path. Combined with a high number of micro-habitats [@Pearson2009], this has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005]. Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided in three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016]. This unparalleled biodiversity is severely threatened by deforestation [@Harper2007; @Vieilledent2013] associated to human activities such as slash-and-burn agriculture and pasture [@Scales2011]. Tropical forest in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large emissions of CO2 in the atmosphere [@Achard2014]. Deforestation threaten species survival by reducing their habitat [@Brooks2002]. Associated to deforestation, forest and habitat fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991]. Edge effects due to fragmentation [@Murcia1995] typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993). Edge effects affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al. 2000). To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined. Detailed information on the extent and distribution of tropical forests and their rate of clearing are essential to estimating threats to biological diversity and carbon emissions.

Accurate maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016]. Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015). In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000). The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s. Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60m (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates. Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example). Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013]. Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015]. However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests, for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015). Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm [@Rakotomalala2015]. In addition, visual interpretation might be subjective and is potentially a source of errors. All this work of image selection and visual interpretation have to be repeated to produce new forest maps in the future.

>>>>

---
title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/biological-conservation.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

@@

Separated from the African continent and the Indian plate about 165 and 85 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path. Combined with a high number of micro-habitats [@Pearson2009], this has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005]. Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided in three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016]. This unparalleled biodiversity is severely threatened by deforestation [@Harper2007; @Vieilledent2013] associated to human activities such as slash-and-burn agriculture and pasture [@Scales2011]. Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO2 emissions in the atmosphere [@Achard2014]. Deforestation threatens species survival by reducing their habitat [@Brooks2002; @Tidd2001]. Associated to deforestation, forest and habitat fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991]. Edge effects due to fragmentation [@Murcia1995].

Dire en quoi la déforestation (perte d'habitat) et la fragmentation (isolation des populations + microclimate on forest edge) affecte la bodiversité. Ainsi, afin de connaitre l'impact de la deforestation sur la biodiversité, il est nécessaire de disposer de cartes forestières spatialisées. Les chiffres de la déforestation et des surfaces forestières seuls, tels qu'ils peuvent être produit par la FAO [@Keenan2015], ne suffisent pas. De même, concernant les émissions de CO2 associées à la déforestation. Les stocks de carbone forestiers varient spatialement [@Vieilledent2016]. Il est donc nécessaires de disposer de cartes de déforestation spatialement explicites.

To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined. Detailed information on the extent and distribution of tropical forests and their rate of clearing are essential to estimating threats to biological diversity and carbon emissions.

Accurate maps of forest-cover and forest-cover change are necessary to monitore loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016].

Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015). In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000). The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s. Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60m (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates. Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example). Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013]. Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015]. However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests, for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015). Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm [@Rakotomalala2015]. In addition, visual interpretation might be subjective and is potentially a source of errors. All this work of image selection and visual interpretation have to be repeated to produce new forest maps in the future.
